---
title: "poskus_00_final"
author: "Luka Močivnik"
date: "3 October 2020"
output: html_document
---

#Intro

This is the latest version of the pipeline used for short DNA samples with high error rates. The default dataset is DAB053_demo01.fastq with tags switched to 16 bp ones.


Setting seed for consistency in random processes/if working on a sample of sequences.
```{r}
set.seed(1)
```


#Tag switch

If necessary this chunck switches tags in the demo data.

```{r, eval=FALSE}
source("./functions/tag_switch_00.R")
system.time(tag_switch())
```


#Error simulation

If necessary this is error simulation.

```{r, eval=FALSE}
source("./functions/error_sim_00.R")
system.time(error_sim())

```


#Data import

Importing sequences, tags, primers, ngsfilter.

```{r}
source("./functions/import_functions_00.R")
demo_data_import(subset_seq = TRUE, subset_size = 10000, input_folder = "./data/error_data/", input_file = "DAB053_demo01.tag16.error10.indel.s34i33d33.fastq")
tag_import()
primer_import()
ngsfilter_import()
```

#Tag and primer search

Searching for primers and tags.

```{r}
if (.Platform$OS.type == "unix") {
  source("./functions/tag_search_00.R")

}else if (.Platform$OS.type == "windows"){
  source("./functions/tag_search_00w.R")

}

system.time(tag_search_par(indels = T))
system.time(primer_search_par(indels = T))

library(data.table)
output = sprintf("./data/tag_primer_hits/hits_primer_f_%s.txt", data_name)
fwrite(x = hits_primer_f, file = output)
output = sprintf("./data/tag_primer_hits/hits_primer_r_%s.txt", data_name)
fwrite(x = hits_primer_r, file = output)
output = sprintf("./data/tag_primer_hits/hits_primer_sum_%s.txt", data_name)
fwrite(x = hits_primer_sum, file = output)
output = sprintf("./data/tag_primer_hits/hits_tag_f_%s.txt", data_name)
fwrite(x = hits_tag_f, file = output)
output = sprintf("./data/tag_primer_hits/hits_tag_r_%s.txt", data_name)
fwrite(x = hits_tag_r, file = output)
output = sprintf("./data/tag_primer_hits/hits_tag_f_filter_%s.txt", data_name)
fwrite(x = hits_tag_f_filter, file = output)
output = sprintf("./data/tag_primer_hits/hits_tag_r_filter_%s.txt", data_name)
fwrite(x = hits_tag_r_filter, file = output)
output = sprintf("./data/tag_primer_hits/hits_tag_sum_%s.txt", data_name)
fwrite(x = hits_tag_sum, file = output)

```

#Primer and tag id

Connecting primers and tags with ngsfilter.

```{r}
if (.Platform$OS.type == "unix") {
  source("./functions/primer_tag_id_00.R")

}else if (.Platform$OS.type == "windows"){
  source("./functions/primer_tag_id_00w.R")

}
system.time(primer_tag_id_parF(indels = T))

output = sprintf("./data/primer_tag_ngs/primer_tag_ngs_%s.txt", data_name)
fwrite(x = primer_tag_ngs, file = output)
```

#Alignment of sample-cluster groups and length correction

```{r}
source("./functions/alignment_00.R")
system.time(alignment())

#source("./functions/alignment_00w.R")
#system.time(alignment())
output = sprintf("./data/fixSeq_alignment_output/fixSeq_%s.txt", data_name)
fwrite(x = fixSeq, file = output)
```

# Clustering and consesnsus

```{r}
source("./functions/clustering_consensus_00.R")
system.time(clustering_par())

output = sprintf("./data/grouped_data_consensus/grouped_data_consensus_%s.txt", data_name)
fwrite(x = grouped_data_consensus_par_mclapply, file = output)
```

#Fishbone allele calling

```{r}
source("./functions/fishbone_prepare_00.R")
fishbone_prepare()

library(fishbone)

out2 <- mclapply(fishbone_input, FUN = callAllele, tbase = mt, verbose = FALSE)#, clean = FALSE))


#fishbone_out <- do.call(rbind, out)
fishbone_out <- do.call(rbind, out2)

output = sprintf("./data/fishbone_out/fishbone_out_%s.txt", data_name)
fwrite(x = fishbone_out, file = output)

```

