---
title: "poskus_00_data_analysis_primer_tag_search_all_files"
author: "Luka MoÄivnik"
date: "10/21/2020"
output: html_document
---

```{r}
library(data.table)
library(ggplot2)
library(dplyr)
library(tidyr)
library(forcats)
library(parallel)
```

#Data import

Reference data import.

```{r}
ref_files = list.files(path = "./data/tag_primer_hits/", pattern = ".+DAB053demo01.tag\\d{1,2}\\.txt$", full.names = TRUE)
ref_files = ref_files[grep(pattern = ".*hits_tag_[f,r]_DAB053.*", x = ref_files, invert = TRUE)]

ref_primer = ref_files[grep(pattern = ".*hits_primer_[f,r]_DAB053.*", x = ref_files)]
ref_tag = ref_files[grep(pattern = ".*hits_tag_[f,r]_filter_DAB053.*", x = ref_files)]

ref_names = list.files(path = "./data/tag_primer_hits/", pattern = ".+DAB053demo01.tag\\d{1,2}\\.txt$")
ref_names = ref_names[grep(pattern = "hits_tag_[f,r]_DAB053.*", x = ref_names, invert = TRUE)]

ref_names_primer = ref_names[grep(pattern = ".*hits_primer_[f,r]_DAB053.*", x = ref_names)]
ref_names_tag = ref_names[grep(pattern = ".*hits_tag_[f,r]_filter_DAB053.*", x = ref_names)]


ref_primer = lapply(ref_primer, fread)
ref_primer = mclapply(ref_primer, as.matrix)
names(ref_primer) = ref_names_primer

ref_tag = lapply(ref_tag, fread)
ref_tag = mclapply(ref_tag, as.matrix)
names(ref_tag) = ref_names_tag

rm(ref_names, ref_names_primer, ref_names_tag)
```

Chekching no error data, if all there are primers and tags found in all sequences.

```{r, eval=F}
lapply(ref_primer, function(x){
  sum((colSums(x)) == 1) == ncol(x)
})

lapply(ref_tag, function(x){
  sum((colSums(x)) == 1) == ncol(x)
})
```

Tag and primer import, for all tags.

```{r}
source("./functions/import_functions_00.R")

primer_import()

tag_import(input_path = "./data/tags_primers/UrsusNGSPrimerslistwithtags.xlsx", input_sheet = "Tag crossbreeding ")

tags8 = tags
tags_u8 = tags_u
tags_uDNA8 = tags_uDNA

tag_import(input_path = "./data/tags_primers/UrsusNGSPrimersCrosbreeding.Ljubljana.18Jul2017_12tag.switch.xlsx", input_sheet = "Tag crossbreeding")

tags12 = tags
tags_u12 = tags_u
tags_uDNA12 = tags_uDNA

tag_import(input_path = "./data/tags_primers/UrsusNGSPrimersCrosbreeding.Ljubljana.18Jul2017_16tag.switch.xlsx", input_sheet = "Tag crossbreeding")

tags16 = tags
tags_u16 = tags_u
tags_uDNA16 = tags_uDNA

rm(tags, tags_u, tags_uDNA)
```

Importing error data.

```{r}
files_primer = "hits_primer_[f,r]_DAB.*s100.*"
err_primer = list.files(path = "./data/tag_primer_hits/", pattern = files_primer, full.names = TRUE)
err_primer_names = list.files(path = "./data/tag_primer_hits/", pattern = files_primer)

files_tag = "hits_tag_[f,r]_filter_DAB.*s100.*"
err_tag = list.files(path = "./data/tag_primer_hits/", pattern = files_tag, full.names = TRUE)
err_tag_names = list.files(path = "./data/tag_primer_hits/", pattern = files_tag)


err_primer = lapply(err_primer, fread)
err_primer = mclapply(err_primer, as.matrix)
names(err_primer) = err_primer_names

err_tag = lapply(err_tag, fread)
err_tag = mclapply(err_tag, as.matrix)
names(err_tag) = err_tag_names

rm(err_primer_names, err_tag_names)
```

Checking number of sequences.

```{r, eval=F}

lapply(err_primer, function(x){
  sum((colSums(x)) == 1)
})
lapply(err_tag, function(x){
  sum((colSums(x)) == 1)
})

```

Finding numbers of individual primers and tags found

```{r}
indPrimerRef = lapply(ref_primer, rowSums)
indTagRef = lapply(ref_tag, rowSums)

indPrimerErr = mclapply(err_primer, rowSums)
indTagErr = mclapply(err_tag, rowSums)

```


Making dataframes for comparisons

```{r, eval=F}
#indPrimerRef$hits_primer_f_DAB053demo01.tag8.txt

# lapply(indPrimerRef, function(x){
#   nameRef = names(x)
# })

indPrimer8df = do.call(rbind, indPrimerErr[grepl(pattern = ".*_f_.*tag8.*", x = names(indPrimerErr))])
indPrimer8df = rbind(indPrimerRef$hits_primer_f_DAB053demo01.tag8.txt, indPrimer8df)

colnames(indPrimer8df) = primers_u
#indPrimer8df

rows = rownames(indPrimer8df)
rows = gsub(pattern = ".*error(\\d{1,2}).*", replacement = "\\1", x = rows)
error = as.integer(rows)

indPrimer8df = cbind(error, indPrimer8df)
indPrimer8df = replace_na(indPrimer8df, 0)

indPrimer8df = indPrimer8df[order(indPrimer8df[,"error"]),]
indPrimer8df
```

Defining function for making tables for comparing found primer and tag numbers

```{r}
numComparison = function(primerTagRef, primerTagErr){
  out = lapply(1:length(primerTagRef), function(x, list, listErr){
  
  nameList = names(list[x])
  nameNum = gsub(pattern = ".*tag(\\d{1,2}).*", replacement = "\\1", x = nameList)
  nameDir = gsub(pattern = ".*_([f,r])_.*", replacement = "\\1", x = nameList)
  searchPattern = sprintf(".*_%s_.*tag%s.*", nameDir, nameNum)
  indPrimerdf = do.call(rbind, listErr[grepl(pattern = searchPattern, x = names(listErr))])
  indPrimerdf = rbind(list[[x]], indPrimerdf)
  
  tagPrimer = gsub(pattern = "hits_(.*)_[f,r]_.*", replacement = "\\1", x = nameList)
  if (tagPrimer == "primer") {
    colnames(indPrimerdf) = primers_u
  }else if(tagPrimer == "tag" & nameNum == "8"){
    colnames(indPrimerdf) = tags_u8
  }else if(tagPrimer == "tag" & nameNum == "12"){
    colnames(indPrimerdf) = tags_u12
  }else if(tagPrimer == "tag" & nameNum == "16"){
    colnames(indPrimerdf) = tags_u16
  }
  
  rows = rownames(indPrimerdf)
  rows = gsub(pattern = ".*error(\\d{1,2}).*", replacement = "\\1", x = rows)
  error = as.integer(rows)
  
  indPrimerdf = cbind(error, indPrimerdf)
  indPrimerdf = replace_na(indPrimerdf, 0)
  indPrimerdf = indPrimerdf[order(indPrimerdf[,"error"]),]

  direction = rep(nameDir, times = nrow(indPrimerdf))
  indPrimerdf = cbind(direction, indPrimerdf)
  
  rownames(indPrimerdf)[1] = nameList
  
  indPrimerdf
    
}, list = primerTagRef, listErr = primerTagErr)
  outnames = names(primerTagRef)
  names(out) = outnames
  return(out)

  
}
```

Using the finction

```{r}
primerNumComparison = numComparison(primerTagRef = indPrimerRef, primerTagErr = indPrimerErr)
tagNumComparison = numComparison(primerTagRef = indTagRef, primerTagErr = indTagErr)
```

Graph for number comparison.

```{r}
graphNumComparison = function(data){
  
  lapply(data, function(x){
    df = as.data.frame(x)
    dfNames = colnames(df)[3:ncol(df)]
    
    fdRowsnames = rownames(df)[1]
    fdRowsnames = gsub(pattern = "_", replacement = " ", x = fdRowsnames)
    fdRowsnames = gsub(pattern = "(.*\\.tag\\d{1,2}).*", replacement = "\\1", x = fdRowsnames)
    fdRowsnames = gsub(pattern = "\\.", replacement = " ", x = fdRowsnames)
    
    dfDraw = df %>%
      gather(dfNames, key = primer, value = number) %>%
      mutate(error = as.factor(as.integer(as.character(error))))
    
    ggplot(dfDraw, aes(fct_inorder(primer), as.integer(number), color = error)) +
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
      ggtitle(fdRowsnames) +
      geom_point()
  })

}

```

```{r}
graphNumComparison(primerNumComparison)
graphNumComparison(tagNumComparison)
```

Finding rercentage of found sequences

```{r}

percData = function(data, zeroRm = TRUE){
  
  data_percentage = lapply(data, function(x, rmZero){
   
    df = as.data.frame(x, stringsAsFactors = FALSE)
    dfNum = as.data.frame(x[,3:ncol(x)], stringsAsFactors = FALSE)
    dfNum = sapply(dfNum, as.integer)
    
    dfPerc = sapply(1:nrow(dfNum), function(y, df){

      perc = round((df[y,]/df[1,]), 3)
      perc

    }, df = dfNum)
  
    dfPerc = t(dfPerc)
    error = as.integer(df$error)
    dfPerc = cbind(error, dfPerc)
    
    if (rmZero == TRUE) {
      dfPerc = dfPerc[2:nrow(dfPerc),]
      rowNamesPerc = rownames(x)[2:nrow(x)]
      rownames(dfPerc) = rowNamesPerc
    }else{
      rowNamesPerc = rownames(x)
      rownames(dfPerc) = rowNamesPerc
    }
    dfPerc
    
  }, rmZero = zeroRm)
  
  namesDperc = names(data)
  names(data_percentage) = namesDperc
  
  return(data_percentage)
}

```


```{r}
primerFoundPerc = percData(data = primerNumComparison)
tagFoundPerc = percData(data = tagNumComparison)
```

Graphs

```{r}
drawPerc = function(data){
  
  datanames = names(data)
  lapply(data, function(x){
    
    df = as.data.frame(x, stringsAsFactors = FALSE)
    nameX = colnames(x)[2:ncol(x)]
    
    drawDf = df %>%
      gather(nameX, key = primer, value = percentage_found) %>%
      na_if(Inf)
    
    fdRowsnames = rownames(df)[1]
    fdRowsnames = gsub(pattern = "_", replacement = " ", x = fdRowsnames)
    fdRowsnames = gsub(pattern = "(.*\\.tag\\d{1,2}).*", replacement = "\\1", x = fdRowsnames)
    fdRowsnames = gsub(pattern = "\\.", replacement = " ", x = fdRowsnames)
    
    ggplot(drawDf, aes(fct_inorder(primer), percentage_found, color = as.factor(error))) +
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
      ggtitle(fdRowsnames) +
      geom_point()
    
  })
}

```

```{r}
drawPerc(primerFoundPerc)
drawPerc(tagFoundPerc)
```

Estimating number of mismatches.

```{r}
diffData = function(data, zeroRm = TRUE){
  diffout = lapply(data, function(x, rmZero){
    
    df = as.data.frame(x, stringsAsFactors = FALSE)
    dfNum = as.data.frame(x[,3:ncol(x)], stringsAsFactors = FALSE)
    dfNum = sapply(dfNum, as.integer)
    
    dfDiff = sapply(1:nrow(dfNum), function(y, df){
      diff = df[y,]-df[1,]
      diff
    }, df = dfNum)
    
    dfDiff = t(dfDiff)
    
    dfDiff = apply(dfDiff, c(1,2), function(y){
      if (y < 0) {
        y = NA
      }else{y = y}
    })
    
    error = as.integer(df$error)
    dfDiff = cbind(error, dfDiff)
    
    if (rmZero == TRUE) {
      dfDiff = dfDiff[2:nrow(dfDiff),]
      rowNamesDiff = rownames(x)[2:nrow(x)]
      rownames(dfDiff) = rowNamesDiff
    }else{
      rowNamesDiff = rownames(x)
      rownames(dfDiff) = rowNamesDiff
    }
    
    
    dfDiff
    
  }, rmZero = zeroRm)
  
  return(diffout)
}

```


```{r}
diffPrimer = diffData(data = primerNumComparison)
diffTag = diffData(data = tagNumComparison)
```

